---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# `allofus` R Package

The goal of the `allofus` R package is to streamline the use of R within the [All of Us Researcher Workbench](https://www.researchallofus.org/data-tools/workbench/).

*The `allofus` R package is a not affiliated with or endorsed by the All of Us Research Program.*

The `allofus` package helps researchers query the database and build transparent and reproducible analysed pipeline. The package allows researchers to easily use other database packages, such as `dbplyr` and `DBI`, with the All of Us database. The package also includes functions to help researchers manage storage in their workspace and bucket, as well as take advantage of tools developed for [OMOP CDM](https://www.researchallofus.org/faq/what-is-omop/) data.

<!-- badges: start -->
[![R-CMD-check](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/allofus)](https://CRAN.R-project.org/package=allofus)
[![](http://cranlogs.r-pkg.org/badges/grand-total/allofus)](https://cran.r-project.org/package=allofus)
[![DOI](https://zenodo.org/badge/659848534.svg)](https://zenodo.org/doi/10.5281/zenodo.10420610)
<!-- badges: end -->

### Installation

Install the released version of `allofus` from CRAN:

```{r}
install.packages("allofus")
```

Install the development version from Github:

```{r}
install.packages("pak")
pak::pak("roux-ohdsi/allofus")
```

On the new RStudio interface on the workbench, you will need to manually specify
the CRAN mirror to be able to download *any* recently updated packages. 

```{r}
# specify the mirror directly
install.packages("allofus", repos = 'https://cloud.r-project.org')

# OR set the mirror as an option at the top of your script
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Github development versions may requires using the remotes package
install.packages("remotes")
remotes::install_github("roux-ohdsi/allofus", repos = 'https://cloud.r-project.org')
```

### Use

Read through the [getting started](https://roux-ohdsi.github.io/allofus/articles/allofus.html) vignette to learn how to use the package.

### Citation

Please cite the `allofus` package as:
```{r, results = "asis", echo = FALSE, eval = TRUE}
print(citation("allofus"), style = "html")
```
or with
```{r}
citation("allofus")
```

### Bugs

Please leave us comments, requests, and report bugs using the "Issues" tab on github located here: https://github.com/roux-ohdsi/allofus/issues.
