---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# `allofus` R Package

<!-- badges: start -->
[![R-CMD-check](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/allofus)](https://CRAN.R-project.org/package=allofus)
[![](http://cranlogs.r-pkg.org/badges/grand-total/allofus)](https://cran.r-project.org/package=allofus)
[![](https://img.shields.io/badge/DOI-10.1093%2Fjamia%2Focae198-007ec6)](https://www.doi.org/10.1093/jamia/ocae198)

<!-- badges: end -->

The goal of the `allofus` R package is to streamline the use of R within the [All of Us Researcher Workbench](https://www.researchallofus.org/data-tools/workbench/). It has 4 primary goals:

1. Facilitate the use of popular `tidyverse` ecosystem of R packages on the Researcher Workbench
2. Help researchers more efficiently and accurately extract and synthesize survey data and EHR data
3. Increase the interoperability between tools created by the Observational Health Data Sciences and Informatics community (OHDSI) for the [OMOP CDM](https://www.researchallofus.org/faq/what-is-omop/)) and the Researcher Workbench
4. Make connecting to the database and managing files simple 

*The `allofus` R package was developed by Louisa Smith and Rob Cavanaugh at [Northeastern University](https://ohdsi.northeastern.edu) and is not affiliated with or endorsed by the All of Us Research Program.*

### Installation

Install the released version of `allofus` from CRAN:

```{r}
install.packages("allofus")
```

Install the development version from Github:

```{r}
install.packages("pak")
pak::pak("roux-ohdsi/allofus")
```

On the new RStudio interface on the workbench, you will need to manually specify
the CRAN mirror to be able to download *any* recently updated packages. 

```{r}
# specify the mirror directly
install.packages("allofus", repos = "https://cloud.r-project.org")

# OR set the mirror as an option at the top of your script
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Github development versions may requires using the remotes package
install.packages("remotes")
remotes::install_github("roux-ohdsi/allofus", repos = "https://cloud.r-project.org")
```

### Use

Read through the [getting started](https://roux-ohdsi.github.io/allofus/vignettes/allofus.html) vignette to learn how to use the package.

### Citation

Please cite the `allofus` package as:
```{r, results = "asis", echo = FALSE, eval = TRUE}
print(citation("allofus"), style = "html")
```
or with
```{r}
citation("allofus")
```

We also encourage you to reference the specific version of the package you use for an analysis. You can look this up with
```{r}
packageVersion("allofus")
```

### Bugs

Please leave us comments, requests, and report bugs using the "Issues" tab on github.
