---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# `allofus` R Package

<!-- badges: start -->
[![R-CMD-check](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/roux-ohdsi/allofus/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/allofus)](https://CRAN.R-project.org/package=allofus)
[![](http://cranlogs.r-pkg.org/badges/grand-total/allofus)](https://cran.r-project.org/package=allofus)
[![](https://img.shields.io/badge/DOI-10.1093%2Fjamia%2Focae198-007ec6)](https://www.doi.org/10.1093/jamia/ocae198)
<!-- badges: end -->


The goal of the `allofus` R package is to streamline the use of R within the [All of Us Researcher Workbench](https://www.researchallofus.org/data-tools/workbench/). It has 4 primary goals:

1. Facilitate the use of popular `tidyverse` ecosystem of R packages on the Researcher Workbench
2. Help researchers more efficiently and accurately extract and synthesize survey data and EHR data
3. Increase the interoperability between tools created by the Observational Health Data Sciences and Informatics community (OHDSI) for the [OMOP CDM](https://www.researchallofus.org/faq/what-is-omop/)) and the Researcher Workbench
4. Make connecting to the database and managing files simple

*The `allofus` R package was developed by Louisa Smith and Rob Cavanaugh at [Northeastern University](https://ohdsi.northeastern.edu) and is not affiliated with or endorsed by the All of Us Research Program.*

### Installation

Install the released version of `allofus` from CRAN:

```{r}
install.packages("allofus")
```

Install the development version from Github:

```{r}
install.packages("pak")
pak::pak("roux-ohdsi/allofus")
```

### Use

Read through the [getting started](https://roux-ohdsi.github.io/allofus/vignettes/allofus.html) vignette to learn how to use the package.

### A note on "Workbench 2.0"

The All of Us Researcher Workbench has migrated to a new, Verily-based platform ("Workbench 2.0"). A few things work differently there compared to the  original workbench:

- **Buckets aren't created automatically.** Workbench 2.0 workspaces no longer come with a `WORKSPACE_BUCKET` set up for you. Run `aou_create_bucket()` once per workspace to create one (or `aou_create_bucket(temporary = TRUE)` for a bucket that auto-deletes its contents after a couple of weeks) before using `aou_ls_bucket()` and related functions.
- **Environment variables are resolved behind the scenes.** `WORKSPACE_CDR` and `GOOGLE_PROJECT` are no longer injected as OS environment variables on workbench 2.0. `allofus` resolves them automatically via the `wb` CLI when the package loads, and caches the result to `~/.aou-env` so subsequent sessions don't need to re-resolve them.

This was last tested on September 2, 2026, in a workbench 2.0 workspace using CDR version 9 and RStudio, with:

```{r}
install.packages("pak")
pak::pak("roux-ohdsi/allofus")
library(allofus)
con <- aou_connect()
```

Workspaces are configured differently as the workbench changes, so please let us know (via the "Issues" tab) if a new workspace doesn't work with this code.

### Citation

Please cite the `allofus` package as:
```{r, results = "asis", echo = FALSE, eval = TRUE}
print(citation("allofus"), style = "html")
```
or with
```{r}
citation("allofus")
```

*Note: A pre-print can be found here: <https://doi.org/10.1101/2024.04.10.24305611>, or feel free to contact us if you can't access the article.*

We also encourage you to reference the specific version of the package you use for an analysis. You can look this up with
```{r}
packageVersion("allofus")
```

### Bugs

Please leave us comments, requests, and report bugs using the "Issues" tab on github.
