---
title: "Searchable Codebook for AllofUs Surveys"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Searchable Codebook for AllofUs Surveys}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

<style>
.row>main {
max-width: 100vw !important;
width: 94vw !important;
}
.container .template_article {
max-width: 94vw !important;
}
.container .row {
justify-content: space-evenly;
}
</style>
```{r setup, include = FALSE}
library(allofus)
```

This table consists of mapped rows from the [publicly available AllofUs Survey Codebook](https://docs.google.com/spreadsheets/d/1Ey8MScRYZ9QyS4izVYScISLMb62QEhSM-ErbG27dNtw/edit#gid=1832128489)
to the AllofUs PPI Vocabulary available on [Athena](https://athena.ohdsi.org/search-terms/start). A small number of rows did not
match between the codebook and the Athena PPI Vocabulary. The code to generate a full table can be found [here](https://github.com/roux-ohdsi/allofus/blob/main/data-raw/pkg_data.R). It can also be accessed in R using `allofus::aou_codebook`. 

More information on the All of Us Surveys is available at the All of Us [Survey Explorer](https://www.researchallofus.org/data-tools/survey-explorer/). PDF versions of the individual surveys are linked in the table.

```{r, echo = FALSE, out.width = "100%"}
cols = c("concept_code", "concept_id", "concept_name", 
         # "concept_class_id",
         "form_name", "field_type", "field_label", "choices")

df = allofus::aou_codebook
df$concept_id = paste0('<a href=https://athena.ohdsi.org/search-terms/terms/',df$concept_id,' target="_blank">',df$concept_id,'</a>')
df$form_name = paste0('<a href=', df$link, ' target="_blank">',df$form_name,'</a>')
df$form_name = as.factor(df$form_name)
# df$concept_class_id = as.factor(df$concept_class_id)
df$field_type = as.factor(df$field_type)

df = df[,cols]

DT::datatable(df,
              rownames = FALSE,
              filter = 'top',
              plugins = "ellipsis",
              extensions = 'Buttons',
              escape = FALSE,
              options = list(
                #autoWidth = TRUE,
                hover = TRUE,
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 20,
                # scrollX = TRUE,
                # scrollCollapse = TRUE,
                columnDefs = list(list(
                  # targets = c(2,6,7),
                  targets = c(2, 5, 6),
                  render = DT::JS("$.fn.dataTable.render.ellipsis( 20, false )")
                ))
              )) |>
  DT::formatStyle(table = _, columns = cols, fontSize = '80%')
```

## Codebook for health history surveys

```{r}
# code from https://laustep.github.io/stlahblog/posts/DT_childTables.html
library(tidyverse)
library(DT)
NestedData <- function(dat, children){
  stopifnot(length(children) == nrow(dat))
  g <- function(d){
    if(is.data.frame(d)){
      purrr::transpose(d)
    }else{
      purrr::transpose(NestedData(d[[1]], children = d$children))
    }
  }
  subdats <- lapply(children, g)
  oplus <- ifelse(lengths(subdats), "&oplus;", "") 
  cbind(" " = oplus, dat, "_details" = I(subdats), 
        stringsAsFactors = FALSE)
}

nested_data <- allofus::aou_health_history %>% 
  select(-c(concept_id_question, form_name)) %>% 
  distinct() %>% 
  rename(concept_id = concept_id_specific) %>%
  mutate(relative = fct_relevel(relative, "self", "daughter", "son", "sibling", "mother", "father", "grandparent")) %>% 
  arrange(category, relative) %>% 
  mutate(relative = as.character(relative)) %>% 
  nest(data = -c(question, condition, concept_id_overall, category))

dat0 <- select(nested_data, category,
               question, condition)
  
dat1 <- select(nested_data, data) %>% 
  pluck(1)

Dat <- NestedData(dat = dat0, children = dat1)
## whether to show row names
rowNames = FALSE
colIdx <- as.integer(rowNames)

## the callback
parentRows <- which(Dat[,1] != "")

callback <- JS(
  sprintf("var parentRows = [%s];", toString(parentRows-1)),
  sprintf("var j0 = %d;", colIdx),
  "var nrows = table.rows().count();",
  "for(let i = 0; i < nrows; ++i){",
  "  var $cell = table.cell(i,j0).nodes().to$();",
  "  if(parentRows.indexOf(i) > -1){",
  "    $cell.css({cursor: 'pointer'});",
  "  }else{",
  "    $cell.removeClass('details-control');",
  "  }",
  "}",
  "",
  "// --- make the table header of the nested table --- //",
  "var formatHeader = function(d, childId){",
  "  if(d !== null){",
  "    var html = ", 
  "      '<table class=\"display compact hover\" ' + ",
  "      'style=\"padding-left: 30px;\" id=\"' + childId + ", 
  "      '\"><thead><tr>';",
  "    var data = d[d.length-1] || d._details;",
  "    for(let key in data[0]){",
  "      html += '<th>' + key + '</th>';",
  "    }",
  "    html += '</tr></thead></table>'",
  "    return html;",
  "  } else {",
  "    return '';",
  "  }",
  "};",
  "",
  "// --- row callback to style rows of child tables --- //",
  "var rowCallback = function(row, dat, displayNum, index){",
  "  if($(row).hasClass('odd')){",
  "    $(row).css('background-color', 'papayawhip');",
  "    $(row).hover(function(){",
  "      $(this).css('background-color', '#E6FF99');",
  "    }, function(){",
  "      $(this).css('background-color', 'papayawhip');",
  "    });",
  "  } else {",
  "    $(row).css('background-color', 'lemonchiffon');",
  "    $(row).hover(function(){",
  "      $(this).css('background-color', '#DDFF75');",
  "    }, function(){",
  "      $(this).css('background-color', 'lemonchiffon');",
  "    });",
  "  }",
  "};",
  "",
  "// --- header callback to style header of child tables --- //",
  "var headerCallback = function(thead, data, start, end, display){",
  "  $('th', thead).css({",
  "    'border-top': '3px solid indigo',", 
  "    'color': 'indigo',",
  "    'background-color': '#fadadd'",
  "  });",
  "};",
  "",
  "// --- make the datatable --- //",
  "var formatDatatable = function(d, childId){",
  "  var data = d[d.length-1] || d._details;",
  "  var colNames = Object.keys(data[0]);",
  "  var columns = colNames.map(function(x){",
  "    return {data: x.replace(/\\./g, '\\\\\\.'), title: x};",
  "  });",
  "  var id = 'table#' + childId;",
  "  if(colNames.indexOf('_details') === -1){",
  "    var subtable = $(id).DataTable({",
  "      'data': data,",
  "      'columns': columns,",
  "      'autoWidth': true,",
  "      'deferRender': true,",
  "      'info': false,",
  "      'lengthChange': false,",
  "      'ordering': data.length > 1,",
  "      'order': [],",
  "      'paging': false,",
  "      'scrollX': false,",
  "      'scrollY': false,",
  "      'searching': false,",
  "      'sortClasses': false,",
  "      'rowCallback': rowCallback,",
  "      'headerCallback': headerCallback,",
  "      'columnDefs': [{targets: '_all', className: 'dt-center'}]",
  "    });",
  "  } else {",
  "    var subtable = $(id).DataTable({",
  "      'data': data,",
  "      'columns': columns,",
  "      'autoWidth': true,",
  "      'deferRender': true,",
  "      'info': false,",
  "      'lengthChange': false,",
  "      'ordering': data.length > 1,",
  "      'order': [],",
  "      'paging': false,",
  "      'scrollX': false,",
  "      'scrollY': false,",
  "      'searching': false,",
  "      'sortClasses': false,",
  "      'rowCallback': rowCallback,",
  "      'headerCallback': headerCallback,",
  "      'columnDefs': [", 
  "        {targets: -1, visible: false},", 
  "        {targets: 0, orderable: false, className: 'details-control'},", 
  "        {targets: '_all', className: 'dt-center'}",
  "      ]",
  "    }).column(0).nodes().to$().css({cursor: 'pointer'});",
  "  }",
  "};",
  "",
  "// --- display the child table on click --- //",
  "// array to store id's of already created child tables",
  "var children = [];", 
  "table.on('click', 'td.details-control', function(){",
  "  var tbl = $(this).closest('table'),",
  "      tblId = tbl.attr('id'),",
  "      td = $(this),",
  "      row = $(tbl).DataTable().row(td.closest('tr')),",
  "      rowIdx = row.index();",
  "  if(row.child.isShown()){",
  "    row.child.hide();",
  "    td.html('&oplus;');",
  "  } else {",
  "    var childId = tblId + '-child-' + rowIdx;",
  "    if(children.indexOf(childId) === -1){", 
  "      // this child has not been created yet",
  "      children.push(childId);",
  "      row.child(formatHeader(row.data(), childId)).show();",
  "      td.html('&CircleMinus;');",
  "      formatDatatable(row.data(), childId, rowIdx);",
  "    }else{",
  "      // this child has already been created",
  "      row.child(true);",
  "      td.html('&CircleMinus;');",
  "    }",
  "  }",
  "});")

datatable(
  Dat, 
  callback = callback, rownames = rowNames, escape = -colIdx-1,
  options = list(
    paging = FALSE,
    searching = FALSE,
    columnDefs = list(
      list(
        visible = FALSE, 
        targets = ncol(Dat)-1+colIdx
      ),
      list(
        orderable = FALSE, 
        className = "details-control", 
        targets = colIdx
      ),
      list(
        className = "dt-center", 
        targets = "_all"
      )
    )
  )
)
```
