---
title: "Searchable Codebook for AllofUs Surveys"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Searchable Codebook for AllofUs Surveys}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<style>
.row>main {
    max-width: 100vw !important;
    width: 94vw !important;
}
.container .template_article {
    max-width: 94vw !important;
}
.container .row {
    justify-content: space-evenly;
}
</style>
```{r setup, echo = FALSE}
library(allofus)
```

This table consists of mapped rows from the [publicly available AllofUs Survey Codebook](https://docs.google.com/spreadsheets/d/1Ey8MScRYZ9QyS4izVYScISLMb62QEhSM-ErbG27dNtw/edit#gid=1832128489)
to the AllofUs PPI Vocabulary available on [Athena](https://athena.ohdsi.org/search-terms/start). A small number of rows did not
match between the codebook and the Athena PPI Vocabulary. The code to generate a full table can be found [here](https://github.com/roux-ohdsi/allofus/blob/main/data-raw/pkg_data.R). It can also be accessed in R using `allofus::aou_codebook`. 

More information on the All of Us Surveys is available at the All of Us [Survey Explorer](https://www.researchallofus.org/data-tools/survey-explorer/). PDF versions of the individual surveys are linked in the table.

```{r, echo = FALSE, out.width = "100%"}
   cols = c("concept_code", "concept_id", "concept_name", "concept_class_id",
            "form_name", "field_type", "field_label", "choices")

  df = allofus::aou_codebook
  df$concept_id = paste0('<a href=https://athena.ohdsi.org/search-terms/terms/',df$concept_id,' target="_blank">',df$concept_id,'</a>')
  df$form_name = paste0('<a href=', df$link, ' target="_blank">',df$form_name,'</a>')
  df$form_name = as.factor(df$form_name)
  df$concept_class_id = as.factor(df$concept_class_id)
  df$field_type = as.factor(df$field_type)
  
  df = df[,cols]

    DT::datatable(df,
                  rownames = FALSE,
                  filter = 'top',
                  plugins = "ellipsis",
                  extensions = 'Buttons',
                  escape = FALSE,
                  options = list(
                    #autoWidth = TRUE,
                    hover = TRUE,
                    dom = 'Bfrtip',
                    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                    pageLength = 20,
                    # scrollX = TRUE,
                    # scrollCollapse = TRUE,
                    columnDefs = list(list(
                      targets = c(2,6,7),
                      render = DT::JS("$.fn.dataTable.render.ellipsis( 20, false )")
                    ))
                  )) |>
                    DT::formatStyle(table = _, columns = cols, fontSize = '80%')
```




