---
title: "Managing files on the workbench"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The `allofus` package includes several functions designed to help you manage and transfer files between your personal workspace and a shared bucket in Google BigQuery. Understanding the difference between these two storage locations is crucial: 

- **Workspace**: This is your personal storage area. Files here are accessible only to you and will be deleted if you delete your environment. 
- **Bucket** : This is a shared storage space. Files in the bucket can be accessed by collaborators on your project and are used for permanent storage.

## Listing Files
### In Your Workspace

Use `aou_ls_workspace()` to list files in your workspace. This function is handy for quickly checking which files you have stored locally.

```{r}
aou_ls_workspace()
```

### In Your Bucket

Similarly, `aou_ls_bucket()` lists files in your bucket. This function can be used to view files that you or your collaborators have saved for shared access.

```{r}
aou_ls_bucket()
```

You can also use the `pattern` argument with these functions to filter the listed files based on a naming pattern.

```{r}
aou_ls_workspace(pattern = "*.csv")
aou_ls_bucket(pattern = "project_*.csv")
```


## Transferring Files

These functions are used in conjunction with R's reading and writing functions. You can store any type of data in both the workspace and the bucket.

### From Workspace to Bucket

Once you've processed or created a file in your workspace, you might want to move it to the bucket for permanent storage or to share it with collaborators. Use `aou_workspace_to_bucket()` for this purpose.

```{r}
write.csv(data, "file1.csv")
aou_workspace_to_bucket("file1.csv")
```


### From Bucket to Workspace

If you need to use a file that a collaborator has saved to the bucket, or if you want to retrieve a file after deleting your environment, use `aou_bucket_to_workspace()`.

```{r}
aou_bucket_to_workspace("file2.csv")
data <- read.csv("file2.csv")
```

## Workflow Example

Here’s a typical workflow using these functions: 

1. **List files in your workspace** : Check what files you currently have. 
2. **Process or create files** : Perform your data analysis or other work in R. 
3. **Save files to your workspace** : Use R’s file handling functions like `write.csv()` or `write.rds()`. 
4. **Transfer to the bucket for sharing or permanent storage** : Use `aou_workspace_to_bucket()`. 
5. **Access shared files from the bucket** : Use `aou_bucket_to_workspace()` to bring files into your workspace as needed.

## Important Considerations 

- **Storage Hygiene** : Regularly clean up your workspace to avoid clutter and manage storage costs. 
- **Backup Important Files** : Use the bucket to backup important files. Files in your workspace (except for notebooks) are not secure against environment deletion.
